---
title: "Sonoma State University - ICLM Analysis"
author: "Rich McGowan"
date: "December 3, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```
###  Initiate the Workspace and Import the 2 Data Sets from CSV Files
###  Produced by a CS Query
```{r Initiate and Import Data, echo=FALSE}
setwd("C:/Users/Richard/OneDrive/Rich Files/SSU/2018/SonomaStateICLM/project")
library(readr)
library(tidyr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
library(kableExtra)
library(tinytex)
#Read in the CSV data produced from Campus Solutions Query and saved in local folder
ICLM_MAIN <- read_csv("ICLM_MAIN.csv",col_names = TRUE)
```
###  Munge the data sets
```{r Munge, echo=TRUE}
ICLM_MAIN$AdmitType <- factor(ICLM_MAIN$AdmitType)
UGRD_MAIN <- data.frame(ICLM_MAIN %>% filter(Career == 'UGRD',UnitsTakenPrgrss > 0))
PBAC_MAIN <- data.frame(ICLM_MAIN %>% filter(Career == 'PBAC',UnitsTakenPrgrss > 0))
#Calculate the number of terms in UGRD_MAIN*it would be the same number for PBAC_MAIN*
numterms <- as.integer(count(distinct(UGRD_MAIN,Term)))
```
### Calculate the FTES by Student College/Major and Course College
```{r Sum FTES Numbers, echo=TRUE}
#Calculate Undergraduate Average FTES based on the distinct total number of terms in the data frame (e.g. summing the Units Taken Profgress dividing by 15 then dividing by 7 regular terms-Fall 2015 thru Fall 2018-gives a 3 year average)
UGRD_FTE_THREEYRAVG <- UGRD_MAIN %>% group_by(StdntCollege,Major,CrseCollege) %>% summarize(FTES= round(sum(UnitsTakenPrgrss)/15)/numterms)
#Group the UGRD_MAIN data frame so that average headcount can be calculated
headcountdf <- distinct(UGRD_MAIN,ID,StdntCollege,Major,Term) %>% group_by(StdntCollege,Major) %>% count(ID)
headcountdf <- headcountdf %>% mutate(hcavg = n/numterms) %>% select(StdntCollege,Major,hcavg)
headcountdf <- headcountdf %>% group_by(StdntCollege,Major) %>% summarize(sum(hcavg))
names(headcountdf)[3] <- "HeadCount"
```
###  Transform the Data Frame into UGRD FTES by Student College/Major by Course/College
```{r Create Pivot, echo=TRUE}
# Transform the Row Values of Course College into Column Values for the Data Frame (using the dcast function)
Major_Course_College_Pivot <- dcast(UGRD_FTE_THREEYRAVG, StdntCollege+Major~CrseCollege)
# Remove the Extended Education CrseCollege Column from the data frame
Major_Course_College_Pivot$`Extended Education` <- NULL
# Replace the NA FTES values with zeros
Major_Course_College_Pivot[is.na(Major_Course_College_Pivot)]=0

```
### Add HeadCount to the Major_Course_College_Pivot data frame
```{r Add HeadCount, ECHO=TRUE}
Major_Course_College_Pivot <- bind_cols(headcountdf,Major_Course_College_Pivot) 
#Drop the superfluous Columns from the data frame
Major_Course_College_Pivot$StdntCollege1 <- NULL
Major_Course_College_Pivot$Major1 <- NULL

```
### Write the Resulting Data Frame Out to a CSV file
```{r Output Dataframe, ECHO=TRUE}
setwd("C:/Users/Richard/OneDrive/Rich Files/SSU/2018/SonomaStateICLM/project")
write.csv(Major_Course_College_Pivot,file = "Major_Course_College_Pivot.csv")
```
### Calculate the FTES by Student College/Enrollment Status and Course College
```{r Sum FTES Numbers 2, echo=TRUE}
#Calculate Undergraduate Average FTES based on the distinct total number of terms in the data frame (e.g. summing the Units Taken Profgress dividing by 15 then dividing by 7 regular terms-Fall 2015 thru Fall 2018-gives a 3 year average)
UGRD_FTE_THREEYRAVG2 <- UGRD_MAIN %>% group_by(StdntCollege,AdmitType,CrseCollege) %>% summarize(FTES= round(sum(UnitsTakenPrgrss)/15)/numterms)
#Group the UGRD_MAIN data frame so that average headcount can be calculated
headcountdf2 <- distinct(UGRD_MAIN,ID,StdntCollege,AdmitType,Term) %>% group_by(StdntCollege,AdmitType) %>% count(ID)
headcountdf2 <- headcountdf2 %>% mutate(hcavg = n/numterms) %>% select(StdntCollege,AdmitType,hcavg)
headcountdf2 <- headcountdf2 %>% group_by(StdntCollege,AdmitType) %>% summarize(sum(hcavg))
names(headcountdf2)[3] <- "HeadCount"
```
###  Transform the Data Frame into UGRD FTES by Student College/Enrollment Status by Course/College
```{r Create Pivot2, echo=TRUE}
# Transform the Row Values of Course College into Column Values for the Data Frame (using the dcast function)
Enrollment_Status_Course_College_Pivot <- dcast(UGRD_FTE_THREEYRAVG2, StdntCollege+AdmitType~CrseCollege)
# Remove the Extended Education CrseCollege Column from the data frame
Enrollment_Status_Course_College_Pivot$`Extended Education` <- NULL
# Replace the NA FTES values with zeros
Enrollment_Status_Course_College_Pivot[is.na(Enrollment_Status_Course_College_Pivot)]=0

```
### Add HeadCount to the Enrollment_Status_Course_College_Pivot data frame
```{r Add HeadCount, ECHO=TRUE}
Enrollment_Status_Course_College_Pivot <- bind_cols(headcountdf2,Enrollment_Status_Course_College_Pivot) 
#Drop the superfluous Columns from the data frame
Enrollment_Status_Course_College_Pivot$StdntCollege1 <- NULL
Enrollment_Status_Course_College_Pivot$AdmitType1 <- NULL

```
### Write the Resulting Data Frames Out to a CSV file
```{r Output Dataframe, ECHO=TRUE}
setwd("C:/Users/Richard/OneDrive/Rich Files/SSU/2018/SonomaStateICLM/project")
write.csv(Major_Course_College_Pivot,file = "Major_Course_College_Pivot.csv")
write.csv(Enrollment_Status_Course_College_Pivot,file = "Enrollment_Status_Course_College_Pivot.csv")

```



###  Output of FTES UGRD Statistics (Average of 3 Years)
```{r Outuput1, echo=TRUE}
kable(Major_Course_College_Pivot) %>% 
kable_styling("striped", full_width = F, position = "center", font_size = 10)

```
###  Output of Undergraduate College Pivot
```{r Outuput2, echo=TRUE}
kable(Major_Course_College_Pivot) %>% 
kable_styling("striped", full_width = F, position = "center", font_size = 10)
```
###  Plot 1 Bar Chart
```{r}
UGRD_FTE_THREEYRAVG %>% ggplot(aes(x=AdmitType,y=TotalUnits,fill=StdntCollege))+
  geom_bar(stat="identity", width=0.5)+
  theme_minimal()
```
### Plot 2 - Bar Chart
```{r}
College_Pivot %>% ggplot(aes(x=StdntCollege,y=TotUnits,fill=CrseCollege)) +
  geom_bar(stat="identity", width=0.75) + coord_flip() +
  theme(legend.title = element_text(colour="blue", size=8, face="bold"))+
  theme(legend.text = element_text(colour="blue", size = 8, face = "bold"))+
  ylab("Total Units")+xlab("Students' Major College")
```

