---
title: "Sonoma State University - ICLM Analysis"
author: "Rich McGowan"
date: "September 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```
###  Initiate the Workspace and Import the 2 Data Sets from Excel Files  
```{r Initiate and Import Data, echo=FALSE}
setwd("C:/Users/Richard/OneDrive/Rich Files/SSU/2018/SonomaStateICLM/project")
library(readxl)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(data.table)
ICLM_ENROLLMENT <- read_excel("ICLM_ENROLLMENT.xlsx")
ICLM_MAIN <- read_excel("ICLM_MAIN.xlsx")
```
###  Munge the data sets
```{r Munge, echo=TRUE}
UGRD_MAIN <- data.table(ICLM_MAIN %>% filter(Career == 'UGRD'))
PBAC_MAIN <- data.table(ICLM_MAIN %>% filter(Career == 'PBAC'))
UGRD_ENROLLMENT <- data.table(ICLM_ENROLLMENT %>% filter(Career == 'UGRD'))
PBAC_ENROLLMENT <- data.table(ICLM_ENROLLMENT %>% filter(Career == 'PBAC'))
#Calculate Undergraduate FTE Statistics
UGRD_FTE_SUMMARY <- UGRD_MAIN %>% group_by(StdntCollege, AdmitType) %>% summarize(HeadCount = n(),TotalUnits = sum(UnitsTakenPrgrs), FTE= round(sum(UnitsTakenPrgrs)/15))
#Sum the Numeric Columns
myNumCols <- which(unlist(lapply(UGRD_FTE_SUMMARY,is.numeric)))
UGRD_FTE_SUMMARY[(nrow(UGRD_FTE_SUMMARY) + 1), myNumCols] <- data.table(colSums(UGRD_FTE_SUMMARY[, myNumCols], na.rm=TRUE))
#Add the TOTAL row label to the last row
UGRD_FTE_SUMMARY[(nrow(UGRD_FTE_SUMMARY)), ncol(UGRD_FTE_SUMMARY)-4] <- 'TOTAL'

```
###  Add the Student's Major College from Main Student Data Frame to the Enrollment Data Frame
```{r Add Student College, echo=TRUE}
UGRD_ALL_COLLEGE <- UGRD_ENROLLMENT %>% mutate(StdntCollege = (UGRD_MAIN$StdntCollege[match(UGRD_ENROLLMENT$ID,UGRD_MAIN$ID)]))
```
###  Pivot Table of Enrolled Units UGRD Major/College by Course/College
```{r Create Pivot, echo=TRUE}
College_Pivot <- UGRD_ALL_COLLEGE %>% select(CrseCollege,StdntCollege,`Sum Progress`) %>%
  group_by(StdntCollege,CrseCollege) %>% 
  summarise(TotUnits = sum(`Sum Progress`))
```
###  Output of Undergraduate FTE Statistics
```{r Outuput1, echo=TRUE}
kable(UGRD_FTE_SUMMARY) %>% 
kable_styling("striped", full_width = F, position = "center", font_size = 10)

```
###  Output of Undergraduate College Pivot
```{r Outuput2, echo=TRUE}
kable(College_Pivot) %>% 
kable_styling("striped", full_width = F, position = "center", font_size = 10)
```
###  Plot Student College vs. Course College Histogram
```{r}
ggplot(data = UGRD_FTE_SUMMARY, aes(x=StdntCollege,y=HeadCount))+
  geom_bar(stat="identity")
```

